---
title: "p8105_hw2_hp2661"
author: "Huizhong Peng"
date: "2024-09-27"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
```

# Problem 1

Import the CSV file and clean the data.

```{r message=FALSE}
dat_NYC = read_csv(
  file = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv"
  ) |> 
  janitor::clean_names() |> 
  select(line, 
         station_name, 
         station_latitude, 
         station_longitude, 
         starts_with("route"), 
         entry, 
         vending, 
         entrance_type, 
         ada) |> 
  mutate(
    entry = ifelse(entry == "YES", TRUE, FALSE), 
    across(route8:route11, as.character)
    ) |> 
  distinct() # remove duplicate rows

skimr::skim(dat_NYC)
```

This dataset contains 19 variables in total, including 15 character variables 
(Line, station name, Route 1-11, vending and entrance type), 2 numeric variables 
(station latitude/longitude), and 2 logical variables (entry and ADA compliance). 
I rename variables, selected the key columns, converted the character variable 
"Entry" to a logical variable, and converted the numeric variables route8-11 to 
the character variables, and removed the duplicate rows.
There is no need to remove NA.
The dimension of the resulting dataset is 684 × 19, and they are tidy as each 
column represents a variable and each row represents one observation.

**Question 1**

```{r results='hide'}
station = dat_NYC |> 
  distinct(station_name, line, .keep_all = TRUE)

nrow(station)
```

There are 465 distinct stations.


**Question 2**

```{r results='hide'}
station |> 
  filter(ada == TRUE) |> 
  nrow()
```

84 stations are ADA compliant. 


**Question 3**

```{r results='hide'}
dat_NYC |> 
  filter(vending == "NO") |> 
  summarize(proportion = mean(entry))
```

38.46% of station entrances / exits without vending allow entrance.


**Question 4**

Reformat data so that route number and route name are distinct variables.

```{r}
dat_NYC_longer = dat_NYC |> 
  pivot_longer(
    cols = starts_with("route"), 
    names_to = "route_number", 
    values_to = "route_name", 
    values_drop_na = TRUE
  ) |> 
  mutate(
    route_number = parse_number(route_number)
  )

head(dat_NYC_longer, 5)
```

```{r results='hide'}
station_A = dat_NYC_longer |>
  distinct(station_name, line, .keep_all = TRUE) |> 
  filter(route_name == "A")

nrow(station_A)

station_A |> 
  filter(ada == TRUE) |> 
  nrow()
```

60 distinct stations serve the A train. 
Of these stations that serve the A train, 17 stations are ADA compliant.


# Problem 2

Import the dataset and clean the sheet.

```{r}
MR = read_excel(
  path = "./202309 Trash Wheel Collection Data.xlsx", 
  sheet = 1, 
  range = "A2:N586"
  ) |> 
  janitor::clean_names() |> 
  filter(!is.na(dumpster)) |> 
  mutate(
    sports_balls = as.integer(round(sports_balls, 0)), 
    source = "MR", 
    year = as.numeric(year)
  )

Professor = read_excel(
  path = "./202309 Trash Wheel Collection Data.xlsx", 
  sheet = 2, 
  range = "A2:M108"
  ) |> 
  janitor::clean_names() |> 
  filter(!is.na(dumpster)) |> 
  mutate(
    source = "Professor"
  )

Gwynnda = read_excel(
  path = "./202309 Trash Wheel Collection Data.xlsx", 
  sheet = 4, 
  range = "A2:L157"
  ) |> 
  janitor::clean_names() |> 
  filter(!is.na(dumpster)) |> 
  mutate(
    source = "Gwynnda"
  )

# combine the datasets

trash_tidy = bind_rows(MR, Professor, Gwynnda)

str(trash_tidy)
```

There are `r nrow(trash_tidy)` observations and `r ncol(trash_tidy)` variables 
including 12 numeric variables (eg., weight_tons, cigarette_butts, homes_powered) 
and 2 character variables (month, source). 

```{r results='hide'}
# total weight of trash collected by Professor Trash Wheel

trash_tidy |> 
  filter(
    source == "Professor"
  ) |> 
  summarize(
    total_weight_prof = sum(weight_tons, na.rm = TRUE)
  )

# total number of cigarette butts collected by Gwynnda in June of 2022

trash_tidy |> 
  filter(
    source == "Gwynnda",
    year == 2022, 
    month == "June"
  ) |> 
  summarize(
    total_ciga_gwy = sum(cigarette_butts, na.rm = TRUE)
  )
```

The total weight of trash collected by Professor Trash Wheel is 216.26 tons.
The total number of cigarette butts collected by Gwynnda in June of 2022 is 18120.


# Problem 3

Import the datasets and clean the data.

```{r message=FALSE, results='hide'}
dat_bakers = read_csv("./gbb_datasets/bakers.csv") |> 
  janitor::clean_names() |> 
  separate(
    baker_name, 
    into = c("baker_first_name", "baker_last_name"), 
    sep = " "
  ) |> 
  rename(baker = baker_first_name)

dat_bakes = read_csv("./gbb_datasets/bakes.csv") |> 
  janitor::clean_names() |> 
  mutate(
    baker = ifelse(baker == "\"Jo\"", "Jo", baker)
  )

dat_results = read_csv("./gbb_datasets/results.csv", skip = 2) |> 
  janitor::clean_names() |> 
  filter(!is.na(result)) |> 
  mutate(
    baker = ifelse(baker == "Joanne", "Jo", baker)
  )

# check 

anti_join(dat_bakers, dat_results)
anti_join(dat_bakers, dat_bakes)

# merge

dat_merge1 = left_join(dat_results, dat_bakes) 
dat_merge2 = left_join(dat_merge1, dat_bakers)

# organize cols

dat_merge_f = dat_merge2 |> 
  distinct() |> 
  select(series, episode, baker, signature_bake, technical, result, 
         show_stopper, everything())
```


```{r}
# summary

skimr::skim(dat_merge_f)

# write csv

write_csv(dat_merge_f, "./gbb_datasets/merge_data.csv")
```

**Cleaning progress:**

- Use reasonable variable names in all datasets.

- In dataset "dat_bakers", the variable "baker_name" had different name and form 
from other two datasets, so separate this variable into two columns "baker_first_name" 
and "baker_last_name", and rename the "baker_first_name" as "baker".

- In dataset "dat_bakes", "Jo" in variable "baker" had a different name form 
from other datasets, so remove the quotation marks around "Jo". 

- In dataset "dat_results", omit the first two useless rows and change "Joanne" 
in column "baker" to "Jo". 


**Briefly Discussion:**

There are 710 rows and 11 columns, including 7 character variables (eg., baker, 
result, show_stopper) and 4 numeric variables (eg., series, episode, technical). 
Variables "signature_bake", "show_stopper" and "technical" have missing values.

**Star bakers / Winners:**

```{r}
star_baker = dat_merge_f |> 
  filter(series >= 5 & series <= 10 & result %in% c("STAR BAKER", "WINNER")) |> 
  select(series, episode, result, baker) |> 
  arrange(series, episode)

star_baker
```

- Nadiya  obtained 3 "star baker" in series 6. 
- Candice obtained 3 "star baker" in series 7. 
- Sophie  obtained 2 "star baker" in series 8. 
- Rahul   obtained 2 "star baker" in series 9. 
- **It is not surprising that they are winners.**\
\
- Nancy obtained 1 "star baker" in series 5. 
- David never obtained "star baker" in series 10. 
- **It is hard to predict that they would be the winners.**

\

**Viewers:**

```{r}
dat_viewers = read_csv("./gbb_datasets/viewers.csv") |> 
  janitor::clean_names()

print(dat_viewers, n = 10)

average_1 = dat_viewers |> 
  pull(series_1) |> 
  mean(na.rm = T)
average_5 = dat_viewers |> 
  pull(series_5) |> 
  mean(na.rm = T)
```

```{r results='hide'}
average_1
average_5
```

The average viewership in Season 1 is 2.77. 
\
The average viewership in Season 5 is 10.04. 



